---
title: "STAA 566 HW2"
author: "Claudia Santillan-Lensink"
date: "2/14/2022"
output: html_document
knit: (function(input_file, encoding) {
  out_dir <- 'docs';
  rmarkdown::render(input_file,
  encoding=encoding,
  output_file=file.path(dirname(input_file), out_dir, 'index.html'))})
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(tidyverse)
library(ggthemes)
library(viridisLite)
library(viridis)
library(dygraphs)
library(xts)
library(lubridate)
```
# Bechdel Score Trends from 1900-2021

This project was completed using a data set of films from 1874 through 2021 with their Bechdel score rating. The Bechdel test, popularized by Alison Bechdel, has a 0-3 rating system. 

A film gets a 1 if it has at least two female characters in it. A film gets a 2 if they speak to each other. The film gets a 3 if they talk to each other about something besides a man. Otherwise, the film gets a 0. "Passing" the Bechdel test requires a score of 3.

The data for this project came from https://www.kaggle.com/alisonyao/movie-bechdel-test-scores which utilized scores from https://bechdeltest.com/

## Read in data and filter out films prior to 1900
```{r readindata}
bechdel <- read.csv("Bechdel.csv")
head(bechdel)
# exclude films prior to 1900
bechdel <- bechdel %>% filter(year > 1899)
#head(bechdel)
bechdel$year <- (as.Date(as.character(bechdel$year), "%Y")) #formatting for xts object compatibility
#head(bechdel)
```

## Summarize and create xts object with summary data by year

```{r summarize}
bechdel_pm <- bechdel %>%
  group_by(year) %>%
  summarise(Mean = mean(rating),
            Median = median(rating),
            Lower = mean(rating)-sd(rating),
            Upper = mean(rating)+sd(rating),
            Pass = sum(rating == 3),
            Total = sum(rating >= 0),
            Pass_percent = Pass / Total)
  
#bechdel_pm

bechdel_pm_xts <- xts::xts(x = bechdel_pm %>% select(Mean, Lower, Upper, Pass_percent), order.by = bechdel_pm$year)


```

## Graph score trend line

```{r graph}
p_bechdel_dyg <- dygraph(bechdel_pm_xts, main = "Average Bechdel Score for Films from 1900-2021") %>%
  dyAxis("x", drawGrid = FALSE) %>%
  dyAxis("y", label = "Average Bechdel Score") %>%
  dyAxis("y2", label = "Percent Pass", independentTicks = TRUE) %>%
  dySeries(c("Lower","Mean", "Upper"),
           label = "Mean",
           color = "green",
           axis = "y") %>%
  dySeries("Pass_percent",
           label = "Pass Percent",
           color = "purple",
           pointSize = NULL,
           strokeWidth =.25,
           axis = "y2")%>% 

  dyRangeSelector()

p_bechdel_dyg
```

